### 2015-06-06 LOI/inc2.tex
no message

\underline{\textbf{Objectives:}}
Exascale scientific discovery will be severely bottlenecked without sufficient
new research into managing and storing the large amounts of data that will be
produced during the simulation, and analyzed for months afterwards.  Our goal
in this project is to address associated I/O and storage challenges in the context 
of current and emerging storage landscapes, and expedite insights into mission 
critical scientific processes. To that end, we will build on the capabilities offered
by ADIOS and DataSpaces that provide I/O abstractions and services, the Sirocco 
peer-to-peer file system under development at Sandia and the object storage 
and annotation expertise of UC Santa Cruz, to develop a application and 
multi-tier storage aware data management solution. Our solutions will provide
novel functionalities and APIs for 
% It will center upon a multi-tier storage-aware approach and incorporate new
%functionalities, extended object metadata and APIs for 
1) specifying, at the application level, data annotations than enable the qualification
 of the relative importance and utility of data objects, and enable them to be 
 mapped at runtime to appropriate storage capabilities across the storage hierarchy;
2) specifying selectable performance/quality/cost tradeoffs from both application and system perspectives; and
3) evaluating and executing these tradeoffs at runtime, in an autonomic manner,
during data placement and movement, using techniques such as application-aware 
data compression (both lossy with specified error bounds and lossless), re-computation, 
reader prioritization, etc.
%2) a selectable performance/quality tradeoff when reading data, and
%the ability to recompute data if needed, and
%3) incorporate reader prioritization for data annotation, placement, and data
%quality when writing data.  
The key challenge in this project is defining and maintaining the metadata
 connecting different object quality and utility levels, and using this metadata 
 to most appropriate manage the placement and access of data object based on 
 user/application intent/contraints and storage system state.
% to best enabling the user with the best possible outcomes 
 %given the user intent and the storage system state. 
The new user APIs required to interact with this rich metadata system will drive effective
use of the entire middleware and storage stack. 
The project is comprised of a team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
UCSC) and storage (UCSC), and connect and coordinate these key storage
components in a seamless fashion.

@@

The I/O and storage system produced from the project will
enable highly flexible data access modes that are common in data analytics. 
\textbf{First}, data re-computation is allowed to re-generate
data with higher accuracy, if the associated computing and I/O cost is lower
than retrieving the target data from a ``slow'' tier. 
\textbf{Second}, exploratory analysis
operations frequently entail an overall data set view followed by targeted data
exploration on extracted features. 
In particular, the ``overview'' access mode gives a quick but approximate
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the high-accuracy data from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. In the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
\textbf{Third}, to ensure available storage for subsequent operations, we will
offer automatic data migration based on user annotations for required data
lifetimes using monitoring and learning techniques. Unlike existing approaches,
this will be tempered both by the user annotations and through learned access
patterns.  While past access patterns may not indicate future access because
the simulation run purpose may have changed, we are focused on scalability
where runs are subsequently larger as the simulation prepares for a capability
run.  By learning from the output and access patterns during this run sequence,
we can accurate decide how to place and organize data for the critical
capability runs. 
\textbf{Fourth}, we anticipate storing multiple data copies, each
compressed in different ways according to the underlying media, some of these
copies will disappear based on storage pressures, but data persistence will be
maintained according to user specifications. Assuming a relatively low latency
cache layer before a tape system, we can offer exploratory data access
reserving pulling data from tape to just the data required. This will save
scientists time and make data stored on tape usable without long delays.

>>>>

\underline{\textbf{Objectives:}}
Exascale scientific discovery will be severely bottlenecked without sufficient
new research into managing and storing the large amounts of data that will be
produced during the simulation, and analyzed for months afterwards.  Our goal
in this project is to address associated I/O and storage challenges in the context 
of current and emerging storage landscapes, and expedite insights into mission 
critical scientific processes. To that end, we will build on the capabilities offered
by ADIOS and DataSpaces that provide I/O abstractions and services, the Sirocco 
peer-to-peer file system under development at Sandia and the object storage 
and annotation expertise of UC Santa Cruz, to develop a application and 
multi-tier storage aware data management solution. Our solutions will provide
novel functionalities and APIs for 
% It will center upon a multi-tier storage-aware approach and incorporate new
%functionalities, extended object metadata and APIs for 
1) specifying, at the application level, data annotations than enable the qualification
 of the relative importance and utility of data objects, and enable them to be 
 mapped at runtime to appropriate storage capabilities across the storage hierarchy;
2) specifying selectable performance/quality/cost tradeoffs from both application and system perspectives; and
3) evaluating and executing these tradeoffs at runtime during data placement and movement, 
in an autonomic manner using models, heuristics and learning and leveraging  
techniques such as application-aware data compression (both lossy with specified error bounds 
and lossless), re-computation,  reader prioritization, etc.
%2) a selectable performance/quality tradeoff when reading data, and
%the ability to recompute data if needed, and
%3) incorporate reader prioritization for data annotation, placement, and data
%quality when writing data.  
The key challenge in this project is defining and maintaining the metadata
 connecting different object quality and utility levels, and using this metadata 
 to most appropriate manage the placement and access of data object based on 
 user/application intent/contraints and storage system state.
% to best enabling the user with the best possible outcomes 
 %given the user intent and the storage system state. 
The new user APIs required to interact with this rich metadata system will drive effective
use of the entire middleware and storage stack. 
The project is comprised of a team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
UCSC) and storage (UCSC), and connect and coordinate these key storage
components in a seamless fashion.

@@

The I/O and storage system produced from the project will
enable highly flexible data access modes that are common in scientific 
applications and associated analysis. 
\textbf{First}, data re-computation is allowed to re-generate
data with higher accuracy, if the associated computing and I/O cost is lower
than retrieving the target data from a ``slow'' tier. 
\textbf{Second}, exploratory analysis
operations frequently entail an overall data set view followed by targeted data
exploration on extracted features. 
In particular, the ``overview'' access mode gives a quick but approximate
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the high-accuracy data from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. In the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
\textbf{Third}, to ensure available storage for subsequent operations, we will
offer automatic data migration.
%based on user annotations for required data and using monitoring and learning techniques. 
 Unlike existing approaches,
this will be driven both by the user annotations and through learned access
patterns in an autonomic and crosslayer manner.  While past access patterns may not indicate future access because
the simulation run purpose may have changed, we are focused on scalability
where runs are subsequently larger as the simulation prepares for a capability
run.  By learning from the output and access patterns during this run sequence,
we can accurate decide how to place and organize data for the critical
capability runs. 
\textbf{Fourth}, we anticipate storing multiple data copies, each
compressed in different ways according to the underlying media, some of these
copies will disappear based on storage pressures, but data persistence will be
maintained according to user specifications. Assuming a relatively low latency
cache layer before a tape system, we can offer exploratory data access
reserving pulling data from tape to just the data required. This will save
scientists time and make data stored on tape usable without long delays.
